---
title: "Install Bioconductor for the first time"
description: |
  Welcome to the Bioconductor community!

  This guide details how to install the latest release of Bioconductor on your
  computer for the first time.
edited: "2022-11-29"
compiled: "`r Sys.Date()`"
output:
  md_document:
    preserve_yaml: true
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, prompt = TRUE, comment = ""
)
```


```{r, include=FALSE}
## Bioconductor version (format: <major>.<minor>)
bioc_major_minor <- BiocManager::version()
```

```{r, include=FALSE}
## R version (format: <major>.<minor>)
r_major_minor_patch <- paste0(R.Version()$major, ".", R.Version()$minor)
r_major_minor <- gsub(
  pattern = "([[:digit:]]+)\\.([[:digit:]]+)\\.([[:digit:]]+)",
  replacement = "\\1.\\2",
  x = r_major_minor_patch
)
```

## Welcome

Welcome to the Bioconductor community!

This guide details how to install the latest release of Bioconductor on your
computer for the first time.

This process is different from updating your Bioconductor installation, which will be described in a separate guide (coming soon).

## Install *R*

The current release of Bioconductor is version `r bioc_major_minor`; it works with <i class="fab fa-r-project"></i> version `r r_major_minor`.
Users of older <i class="fab fa-r-project"></i> versions must update their installation to take advantage of new features and to access packages that have been added to Bioconductor since the last release.

### Windows

- Navigate to the [CRAN][cran-website] website.
- Follow the links 'Download R for Windows' and 'install R for the first time'.
- Click the link 'Download R-x.x.x for Windows' (`x.x.x` being the latest version of <i class="fab fa-r-project"></i>).
- Run the `.exe` file that was just downloaded.

### macOS

- Navigate to the [CRAN][cran-website] website.
- Follow the link 'Download R for macOS'.
- Click on the `.pkg` file to download the latest <i class="fab fa-r-project"></i> version.
- Double click on the downloaded file to install <i class="fab fa-r-project"></i>.

Additionally:

- It is also a good idea to install [XQuartz][xquartz] (needed by some packages).

### Linux

- Navigate to the [CRAN][cran-website] website.
- Next to the link 'Download R for Linux', click on the name of your Linux distribution.
- Follow the instructions on the web page.

The [CRAN][cran-website] web pages provide information to get the most recent version of <i class="fab fa-r-project"></i> for common distributions.
For most distributions, you could use your package manager (e.g., for Debian/Ubuntu run `sudo apt-get install r-base`, and for Fedora sudo `yum install R`), but we don’t recommend this approach as the versions provided by this are usually out of date.

## Install the BiocManager package

The `r BiocStyle::CRANpkg("BiocManager")` package is a convenient tool distributed on the [CRAN][cran-website] repository to install and update Bioconductor packages.

To install the package, type the following in an <i class="fab fa-r-project"></i> console:

```{r, eval=FALSE}
install.packages("BiocManager")
```

## Install the current Bioconductor release

To install the latest version of Bioconductor core packages, type the following in an <i class="fab fa-r-project"></i> console:

```{r, echo=FALSE, results='asis'}
cat(paste0(
    "```r", "\n",
    sprintf('> BiocManager::install(version = "%s")', bioc_major_minor), "\n",
    "```", "\n"
))
```

## Check your installation

To check your installation, restart your <i class="fab fa-r-project"></i> session, and type the following in the <i class="fab fa-r-project"></i> console:

```{r}
BiocManager::version()
```

## Final words

Congratulations!
You are ready to install the latest version of more Bioconductor packages!

<!-- Links -->

[bioc-intro-setup]: https://carpentries-incubator.github.io/bioc-intro/setup.html
[cran-website]: https://cran.r-project.org/
[xquartz]: https://www.xquartz.org/